---
title: "Coding Challenge 6"
author: "Maryam Saeed"
date: "2025-03-27"
output:
  word_document: default
  html_document: default
md_document:
  variant: gfm
---

# Q.No.1

### Writing function and iteration helps to do repetative tasks and save alot of time it allow copy pasting and make the script easy to read and use


# Q.No.2
**To write function**

### To write a function we use function() keyword to perform a task. then function() take the input arguments and return the results by applying results().

**To write a loop**

### in the order to make a loop we use for and put i as a iteration in the data and then print i. In this way same code run for each one



```{r}
### How to write a function in R
F_to_C <- function(f_temp){     # Example of writing a function
  Celcius<- (5*(f_temp -32)/9)
  return(Celcius)
}

F_to_C(80)
```

```{r}
for(i in -10:10){           # writing a for loop function in R
  result<- F_to_C (i)
  print(result)
}
```

### syntax to write code and how codes are returned
#### function(argumnet){(code of operation) return(results)}

# Q.No.3 Read in cities.csv file using a relative path

```{r}
cities <- read.csv("Cities.csv")
head(cities)
```
# Q.No.4

```{r}
haversine_distance <- function(lat1, lon1, lat2, lon2){
 rad.lat1 <- lat1 * pi/180   # converting degrees to radians
 rad.lon1 <- lon1 * pi/180
 rad.lat2<- lat2* pi/180
 rad.lon2<- lon2 * pi/180
 delta_lat<- rad.lat2 - rad.lat1   # Haversine formula
 delta_lon<- rad.lon2 - rad.lon1
  a<- sin(delta_lat/2)^2 + cos(rad.lat1) * sin(delta_lon/2)^2
  c<-2 *asin(sqrt(a))
earth_radius<- 6378137    # Earth radius in kilometer
distance_km<- (earth_radius * c)/1000 # final distance in kilometer
return(distance_km)
}
```

# Q.No.5(a)

```{r}
# filter for newyork city and auburn
library(dplyr)
nyc<- cities %>% filter(city == "New York") 
auburn<- cities %>% filter(city == "Auburn")

# Extracting just latitude and longitude values
lat1<- nyc$lat
lon1<- nyc$long

lat2<- auburn$lat
lon2<- auburn$long

# Putting the inputs in the function
haversine_distance( lat1, lon1, lat2, lon2)
```

# Q.No.5(b)

```{r}
haversine_distance( lat1, lon1, lat2, lon2)
```

# Q.No.6

```{r}
# loops to calculate distance from auburn to each city
for(i in 1:nrow(cities)){
  if(cities$city[1] != "Auburn"){
    lat1<- cities$lat[i]
    lon1<- cities$long[i]
    
    dist<- haversine_distance(lat1, lon1, lat2, lon2)
    print(dist)
  }
}
```


# Work for bonus point

```{r}
# creating a empty data frame to store results
distance_df<- data.frame(city1 = character(),
                         city2 = character(),
                         distance_km = numeric())
# Loop through each city
for(i in 1:nrow(cities)){
  if(cities$city[i] != "Auburn"){
    lat1<- cities$lat[i]
    lon1<- cities$long[i]
# calculating the distance
    dist<- haversine_distance(lat1, lon1, lat2, lon2)
    
# Adding results as new row
    new_row<- data.frame(city1 = cities$city[i],
                         city2 = "Auburn",
                         distance_km = dist)
    
    distance_df<- rbind(distance_df, new_row)
  }
  
}

# first six rows of the data frame
head(distance_df)
```

# Q.No.7 Adding a link to my github

(Github account)[https://github.com/meriumm/Reproducibility_5820.git]

















